OpenMotion: An open source library for attitude estimation.

### 2016-08-04 Article/conclusion.tex
Modification Article, Calibration file + Introduction + Conclusion

notre library est trop bien...

>>>>

In this paper, we presented our open source library \texttt{OpenMotion} and its different functionalities. Our aim is to make it a powerful tool for developing application or new algorithms. Combining this library with our comparative framework \cite{braudcomparison} could be an interesting tool for research. A performance study has been processed too in order to show the capacities of the algorithms present in the library. Thus, the user has access to the algorithm that fit the best to his needs and the open source feature allows him to have a better understanding of the algorithms. For future work, we continue to develop the library and focus on his robustness and adaptability to every type of smartphone. With the developing sensor calibration, better accuracy is expected too.

### 2016-08-04 Article/introduction.tex
Modification Article, Calibration file + Introduction + Conclusion

>>>>

%context + motivation

Since the last decades, it is easily observable that smartphones market has grown drastically. Indeed, this devices propose convenient functionalities for household. On the other hands, the different kind of sensor available and easily accessible make smartphones an interesting tool for research.  Nowadays, some applications where smartphone can be used attract the attention of researchers such as Indoor Navigation\cite{huang2009survey}. Severals solution for this has been proposed with different approaches like using only computer vision technics \cite{desouza2002vision,bonin2008visual} or wireless networks\cite{gu2009survey}. Another approaches is to sense the motion by using some inertial sensors. Among its sensor and according to the smartphone, an Inertial Measure Unit (IMU) is vacant and used when coupling with a magnetometer as an Attitude and Heading Reference System (AHRS). However, an AHRS sensor may be not robust or accurate in certain situation due to unpredictable event or sensor deficiencies. Moreover, for Indoor Localization or Navigation applications, it is crucial to have an algorithm that propose an optimal estimation of the attitude in term of accuracy and computation time requirement. This is the reason why researchers are interested in sensor fusion approaches because it overcomes some problems such as sensor drifting.

Since 1960, severals data fusion has been developed after a seminal framework on data fusion has been initiated by Kalman and  Bucy \cite{kalman_new_1960}. In this paper, the most common family of data fusion algorithm is presented briefly. However, implementing this kind of algorithm require time and resources. Furthermore, the available software of data fusion remains general and may be not adaptable for the inertial case. For this reason, we propose an open source and free library called \texttt{OpenMotion} that contains different family of sensor fusion algorithms designed to fit to the inertial case. For this paper, the different functionalities proposed by the library are presented. In a simulated environment, a performance study of all methods has been processed and the values are shown. To quantify the performances, we determined 2 kind of scores,  $s_1$ representing the accuracy of the method and $s_2$ that takes into account the accuracy and the computation time. The $s_2$ give better visibility of the methods' effectiveness for usage in embedded devices, which is interesting for smartphone application. Furthermore, working only in a simulated environment is not enough to have a good visibility. Moreover, in our context, it is very difficult to obtain a reliable ground truth. For this study, we designed the ground truth  based on computer vision methods.

### 2016-08-04 Article/prez_library.tex
Modification Article, Calibration file + Introduction + Conclusion

\texttt{OpenMotion} is a open source library mainly aimed for real-time attitude estimation. Our approach is based on data fusion technics using only the output of a IMU (composed by a 3D gyroscope,  3D accelerometer and a 3D magnetometer). Among several methods proposed by the library, the user has the possibilities to choose which methods fits the best to his expectation according to its performance. Here the 2 reference frame to remember.

@@

For simplicity reasons, the library admit that all the components of the IMU belong to the same reference frame $\{c\}$ (ref figure \ref{Schema_situation}). It is a cyclopean approximation \cite{ouarti2008multimodal}. Moreover, the user has to interface the sends of data from the device to the library because the library does not take care of it. \texttt{OpenMotion} provide an estimation of the orientation of $\{c\}$ relative to $\{a\}$. The choice of the output type (quaternion, rotation matrix, euler angle) is according to the user needs.

\begin{figure}
\centering
\includegraphics[scale=0.65]{images/Schema_situation.png}
\caption{Definition of the scene with 2 frame:  a fixed frame NED (North East Down) noted $\{a\}$ and a moving frame object noted $\{c\}$}
\label{Schema_situation}
\end{figure}

%
%\begin{lstlisting}[caption=Exemple of utilisation in C++,label=code]
%// om_exemple.cpp
%#include <om.h>
%
%using namespace om;
%
%void main(int argc, char** argv){
%
%	// 
%	SensorFusionManager manager;
%
%	//if calibration is necessary
%	manager.calibration();
%	
%	// while IMU send data
%	while( isData ){
%
%		// set all IMU data at time t
%		manager.setGyroscopeData(gyro_x,gyro_y,gyro_z);	
%		manager.setAccelerometerData(acc_x,acc_y,acc_z);
%		manager.setMagnetometerData(mag_x,mag_y,mag_z);
%		
%		// get different kind of output according to user's choices
%		
%		// quaternion
%		Quaternion q_est = manager.getAttitude<Quaternion>(); 
%		
%		//  rotation matrix
%		Matrix R = manager.getAttitude<Matrix>(); 
%		
%		//  axis angle
%		AxisAngle axisAngle = manager.getAttitude<Matrix>(); 
%		...
%	}
%}
%
%\end{lstlisting}

>>>>

\texttt{OpenMotion} is a open source library mainly aimed for real-time attitude estimation. Our approach is based on data fusion technics using only the output of a IMU (composed by a 3D gyroscope,  3D accelerometer and a 3D magnetometer). Among several sensor fusion algorithms proposed by the library, the user has the possibilities to choose the one that fits the best to his expectation according to its performance.

However, we designed the library also as an academic tool. For this reason, we have chosen to implement for this library the most common family of methods. In previous works \cite{braudcomparison}, we proposed a comparative framework that allow us to observe the performance of this different methods in a quantitative manner. Thus, the user has in his disposition the code and and the performance study. This will help him to have a better understanding of the algorithms' mechanism and behavior. On the other hand, the user can use the library as a developing tool to design and create new sensor fusion algorithm. Moreover, he can compare his work directly with the available methods. For this paper,  2 references frames are used as follow:

@@

For simplicity reasons, the library admit that all the components of the IMU belong to the same reference frame $\{c\}$ (ref figure \ref{Schema_situation}). It is a cyclopean approximation \cite{ouarti2008multimodal}. Moreover, the user has to interface the sends of data from the device to the library because the library does not take care of it. \texttt{OpenMotion} provide an estimation of the orientation of $\{c\}$ relative to $\{a\}$. The choice of the output type (quaternion, rotation matrix, euler angle) is also according to the user needs.

\begin{figure}
\centering
\includegraphics[scale=0.65]{images/Schema_situation.png}
\caption{Definition of the scene with 2 frame:  a fixed frame NED (North East Down) noted $\{a\}$ and a moving frame object noted $\{c\}$. \texttt{OpenMotion} will provide an estimate of the attitude of $\{c\}$ relative to $\{a\}$. }
\label{Schema_situation}
\end{figure}

### 2016-08-04 Article/state_of_the_art.tex
Modification Article, Calibration file + Introduction + Conclusion

The matrix $R(\mathbi{q})$, is the rotation matrix that corresponds to the quaternion $\mathbi{q}$ (equation \ref{quat_to_rot}). The Quatenion Estimator (QUEST) algorithm propose a solution to the Wahba's problem. Let :

\begin{equation}
B = \sum_{i=1}^N a_i\textbf{b}_i \textbf{r}_i^T
\end{equation}

Then, we can rewrite the equation (\ref{QUEST}) as follow

\begin{equation}
J(\mathbi{q}) = \frac{1}{2} \sum_{i=1}^n a_i  - tr(R(\mathbi{q})B^T)
\label{QUEST_bis}
\end{equation}

Due to the fact that the representation of the attitude matrix is a homogenous quadratic function of $\mathbi{q}$, we can say that

\begin{equation}
tr(R(\mathbi{q})B^T) = \mathbi{q}^TK\mathbi{q}
\label{QUEST_ter}
\end{equation}

where $K$ is the symmetric traceless matrix such as

\begin{equation}
K = \begin{pmatrix} S-tr(B)I_3 & \textbf{z} \\ \textbf{z}^T & tr(B)
\end{pmatrix}
\end{equation}

with

\begin{equation}
\left\{\begin{array}{l}
S = B + B^T\\
\textbf{z} = \sum_ia_i\textbf{b}_i\times\textbf{r}_i
 \end{array}
\right.
\end{equation}

The minimization problem from equation (\ref{QUEST_bis}) is now a maximization problem. thus, the solution $\mathbi{q}_{opt}$ is the eingen vector correspond to the highest eigen value of $K$ \cite{markley1999estimate}. The main problem of QUEST is that it does not take into account of all the previous observations performed since the beginning. Then, a REcursive-QUEST algorithm (REQUEST) proposed by Bar-Itzhack\cite{bar-itzhack_request-recursive_1996} manages this problem.  Contrary to other methods, solutions to Wahba's problem do not estimate the gyroscope bias. Which can make it less robust in some cases. There are also other methods based on Wahba's problem like the Davenport's q-method \cite{weighted1971nasa} proposed in 1971 considered more robust than QUEST. Or the method ESOQ proposed by Mortari \cite{mortari1997esoq}  in 1997. However, for this library, we limit our investigation to REQUEST.

@@

With $n$ the number of iteration at our disposal. We use this information (mean error) for our performance study. To differentiate the methods, we assigned  some scores in order to highlight the behaviour of these methods in different configurations defined as:

%\begin{figure}[!h]
%\centering
%\includegraphics[scale=0.40]{images/test_method.png}
%\caption{Algorithm designed to compute attitude estimation methods' performances}
%\label{test_method}
%\end{figure}

@@

\begin{table}[!h]
\begin{tabular}{|c|c|c|}
\hline
Sensor & Bias instability  & Measurement noise\rule[-2pt]{0pt}{10pt} \\
\hline
\hline
 Gyroscope & $0.316\times 10^{-4} $  $\mu$rad.s$^{-\frac{3}{2}}$ & $ 0.316$  $\mu$rad.s$^{-\frac{1}{2}}$ \rule[-1.5pt]{0pt}{13pt}\\
 \hline
Magnetometer & $0.0023$ mGauss$/\sqrt{Hz}$ & 0.05 mGauss  \rule[-1.5pt]{0pt}{13pt}\\
 \hline
Accelerometer &  $0.002$ m.s$^{-2}/\sqrt{Hz} $ & 0.02 m.s$^{-2}$ \rule[-1.5pt]{0pt}{13pt}\\
 \hline
\end{tabular}
\caption{Calibrated IMU data performance specification, Gyrometer, Accelerometer, Magnetometer}
\label{spec_imu}
\end{table}

>>>>

The matrix $R(\mathbi{q})$, is the rotation matrix that corresponds to the quaternion $\mathbi{q}$ (equation \ref{quat_to_rot}). The Quatenion Estimator (QUEST) algorithm propose a solution to the Wahba's problem. The main problem of QUEST is that it does not take into account of all the previous observations performed since the beginning. Then, a REcursive-QUEST algorithm (REQUEST) proposed by Bar-Itzhack\cite{bar-itzhack_request-recursive_1996} manages this problem.  Contrary to other methods, solutions to Wahba's problem do not estimate the gyroscope bias. Which can make it less robust in some cases. There are also other methods based on Wahba's problem like the Davenport's q-method \cite{weighted1971nasa} proposed in 1971 considered more robust than QUEST. Or the method ESOQ proposed by Mortari \cite{mortari1997esoq}  in 1997. However, for this library, we limit our investigation to REQUEST.

@@

With $n$ the number of iteration at our disposal. We use this information (mean error) for our performance study. To differentiate the methods, we assigned  some scores in order to highlight the behavior of these methods in different configurations defined as:

@@

\begin{table}[!h]
\begin{tabular}{|c|c|c|}
\hline
Sensor & Bias instability  & Measurement noise\rule[-2pt]{0pt}{10pt} \\
\hline
\hline
 Gyroscope & $0.316\times 10^{-4} $  $\mu$rad.s$^{-\frac{3}{2}}$ & $ 0.316$  $\mu$rad.s$^{-\frac{1}{2}}$ \rule[-1.5pt]{0pt}{13pt}\\
 \hline
Magnetometer & $0.0023$ mGauss$/\sqrt{Hz}$ & 0.05 mGauss  \rule[-1.5pt]{0pt}{13pt}\\
 \hline
Accelerometer &  $0.002$ m.s$^{-2}/\sqrt{Hz} $ & 0.02 m.s$^{-2}$ \rule[-1.5pt]{0pt}{13pt}\\
 \hline
\end{tabular}
\caption{Calibrated IMU data performance specification, Gyroscope, Accelerometer, Magnetometer}
\label{spec_imu}
\end{table}
